---
title: "Mental Health in Australia"
subtitle: "My subtitle if needed"
author: 
  - Christina Chan-Ying
  - Yan Dumalag
  - Reese Halfyard
thanks: "Code and data are available at: https://github.com/ChristinaChanYing/Australian-Social-Determinants-of-Health"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
nocite: '@*'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(ggplot2)

library(tibble)
library(reshape2)
```

```{r, include=FALSE}
### Grabbing the data ###
female_data <- readr::read_csv(here::here("inputs/data/cleaned_data_females.csv")) |>
   as_tibble() |>
  select(Section, Modules, `Has a mental health condition`, `Does not have a mental health condition`, `Has a long term health condition`, `Does not have a long term health condition`, `Has disability`, `Has no disability`, `Total persons`) |>
  mutate(Gender = "female")

male_data <- readr::read_csv(here::here("inputs/data/cleaned_data_males.csv")) |>
   as_tibble() |>
  select(Section, Modules, `Has a mental health condition`, `Does not have a mental health condition`, `Has a long term health condition`, `Does not have a long term health condition`, `Has disability`, `Has no disability`, `Total persons`)|>
  mutate(Gender = "male")

persons_data <- readr::read_csv(here::here("inputs/data/cleaned_data_persons.csv")) |>
   as_tibble() |>
  select(Section, Modules, `Has a mental health condition`, `Does not have a mental health condition`, `Has a long term health condition`, `Does not have a long term health condition`, `Has disability`, `Has no disability`, `Total persons`)

female_male_data <-
    merge(x = female_data, y = male_data, by = c("Section", "Modules", "Has a mental health condition", "Does not have a mental health condition", "Has a long term health condition", "Does not have a long term health condition", "Has disability", "Has no disability", "Total persons", "Gender"), all = TRUE) |>
  group_by(Gender)

```
# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point). 

# Data

This paper was written in R, a programming language for statistical computing [@citeR] and uses various R packages for data analysis and manipulation. These packages include readxl [@readxl] for reading the .xlsx file and tidyverse [@tidyverse] and dplyr [@dplyr] for handling the data. The data itself was the 2020 Australia General Social Survey: Summary Results – Tables 3: Persons aged 15 years and over, Social Experiences–By Sex and Health Characteristics obtained from the Australian Bureau of Statistics [@ausstats]. Since the data was an .xlsx file and contained three separate tables (Males, Females, and Persons) in a single sheet, we split the table in excel into three separate sheets. We also took the subheading in each table and created another column titled ‘Section’ in excel to make it easier to convert to a .csv file in R and manipulate the data. 

## Source and Methodology

## Questionnaire Details
The questionnaire collected data from approximately 5,300 households around Australia which is a generally good sample size. However, the survey did not cover people who live in “very remote parts of Australia” and “discrete Aboriginal and Torres Strait Islander communities”. 
As such, this exclusion in the survey may have been either done to make the survey distribution or data collection process simpler or as a result of the bias against discrete aboriginals, Torres Strait Islanders, or ruralites in general. Given that this survey was completed online or via a telephone interview, it is unlikely that the first reason is responsible for this exclusion, especially given that there is a history of Australia’s implicit racial bias against aboriginals and Torres Strait Islanders. Coverage of discrete aboriginals and Torres Strait Islanders would have been beneficial for our study since “[h]istorically they have been mistreated by healthcare institutions with inequitable care prevalent within emergency settings” [@aboriginalbias]. 
 
One of the major drawbacks of this survey was that compared to their previous survey, this one had no face-to-face interviews. As such, the answers received may not be as accurate as previous years’ questionnaires since the interviewer would be unable to capture the interviewee’s verbal and non-verbal cues. Additionally, this has caused there to be a “higher than usual proportion of non-response (37.9%)” [@ausstats]. 
 
The questions asked had a good selection of options for the participant to choose from and often included more open-ended options such as “Other (please specify)” or “Not stated [Allow blank response/skip question functionality]” for people who may not fall into the other options. For instance, when asking for gender, the questionnaire had "male", "female", and "other" options for selection. While most of their questions had a good variety of answers to select from, there were some exclusions in the questions they asked. For example, while there are various questions relating to the participant’s marital status, none were asking about their relationship status. This exclusion ignores the people who are in a serious relationship but do not wish to get married. This could lead to the survey missing some important information since being in a serious relationship with another individual often has a similar effect on one’s mental well-being to a married relationship. 
 
The questions were explicit and easy to read and understand. Many of the questions included an example to ensure that the participant understood what the question was asking and that they know how to properly answer it. The questions also had a good flow and order to them. For instance, in “Module 7.2: Contact with Family and Friends (CWF)” they would start by asking “In the last 3 months have you seen family or friends who do not live with you in person?”, then follow up with asking other questions relating to it such as “How often have you seen them in person in the last 3 months?” or “In the last 3 months, have you used any of these other types of contact with family or friends [CWF_WS1]?”. 

# Results

# Discussion

## Stress and Anxiety

In comparison to the feelings of anxiety or stress, 700 females with mental health issues reported “Always” or “Often” to feeling stressed or rushed. This was much higher than males, in which only 300 participants felt consistent stressors or anxiety. Due to societal pressures and expectations, it is extremely likely that women would feel stressors, much more often than men would. According to The Guardian, Australian women encounter stress on a much higher level than men do. This is in regards to finances, family, and physical health (The Guardian Reference). Citizens with mental health conditions already experience difficulties from other day activities and events. The addition of a permanent illness will greatly affect the ways Australian men and women approach daily life. 

There are many other factors that affect Australian citizens' experiences with stress and anxiety. This includes intersectionality of race, gender, ethnicity, and sexual orientation. In the survey, 2,168.4 people without mental illness and 538.9 people with mental illness reported experiencing some form of discrimination in the past 12 months. This justifies the fact that those with mental health conditions are not as likely to experience oppression because of their illness. For example, a white male with mental illness will have different experiences in navigating daily anxiety and stress than a black woman would. Women, LGBTQ+, and BIPOC experience systematic oppression in a white straight male dominated society. Those who do not identify with these groups may experience difficulties due to their mental illness, but will not face discimination because of their identity. 

```{r stressors, fig.cap="Male vs Female Stressors", echo = FALSE}
### Grabbing the stressors ###
stressors <- subset(female_male_data, Section == "Stressors", select = -c(Section))

### Plotting the stressors ###
ggplot(stressors, aes(x = Modules, y = `Total persons`, group = Gender, fill = Gender)) +
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_text(aes(label = `Total persons`), vjust=1.6, color="black",
            position = position_dodge(0.9), size=3.5)+
  theme_minimal() +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 20)) +
  ylab("Total Population") +
  scale_fill_manual(values=c('#d9848e','#84a8d9'))

```

## View on Healthcare
```{r healthcare, fig.cap="Percent of Australians that agreed the healthcare system can be trusted in 2020", echo = FALSE}
### Pie Chart Based on Secondary Research ###
percent_trust <- c(76, 24)
lbls <- c("Agree", "Disagree")
pct <- round(percent_trust/sum(percent_trust)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pie(percent_trust, labels = lbls, col=c("#bce067", "#ffbd58"), main = "The healthcare system can be trusted")

```


In this survey, we understand the differences between men and women who trust the healthcare system with and without mental health conditions. According to Figure 2, this is quite a drastic contrast. 995 women with mental health conditions reported that they have trust among most people, and men sat slightly lower, with 519.7 people. Men without mental illness are at 5,563, whereas women without are 5,343. The key question is why is the amount of trust so much higher in those without mental illness, than those with? Lack of trust in most people equates to common everyday services, including healthcare and treatment. In order to understand the lack of trust that those with mental illness or disabilities have regarding the healthcare system, we must look at history. Up until a few decades ago, the treatment of those with bipolar disorder, depression, anxiety, and other common illnesses, were through a myriad of inhumane theories and practices. In 2006, the Bureau of Justice Statistics reported that “705,600 mentally ill men were incarcerated in the state prison system, and another 78,800 were incarcerated in the federal prison system” [@opentextbc].

Only until recently, considerations and understandings of mental illness in modern society have been considered. However, a history of societal neglect justifies the lack of trust which both male and female Australians with mental illness obtain.

```{r mhc-trust, fig.cap="People that strongly agrees most people can be trusted", echo = FALSE}
### Get trust data on people with mental health conditions ###
has_mhc_trust <- subset(female_male_data, Section == "Feels most people can be trusted" & Modules == "Strongly agree / Somewhat agree") |>
  select(`Has a mental health condition`, Gender) |>
  rename(Population = `Has a mental health condition`) |>
  mutate(`Has a mental health condition` = "Yes")

### Get trust data on people without mental health conditions ###  
no_mhc_trust <- subset(female_male_data, Section == "Feels most people can be trusted" & Modules == "Strongly agree / Somewhat agree") |>
  select(`Does not have a mental health condition`, Gender) |>
  rename(Population = `Does not have a mental health condition`) |>
  mutate(`Has a mental health condition` = "No")

### Merge data from people with and without mental health conditions ###
mhc_trust <- merge(x = has_mhc_trust, y = no_mhc_trust, all = TRUE)

### Plot trust data ###
mhc_trust <- ggplot(mhc_trust, aes(x = `Has a mental health condition`, y = Population, group = Gender, fill = Gender)) +
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_text(aes(label = Population), vjust=1.6, color="black",
            position = position_dodge(0.9), size=3.5)+
  theme_minimal()
mhc_trust + scale_fill_manual(values=c('#d9848e','#84a8d9'))

```

```{r trust, fig.cap = "Feels the healthcare system can be trusted", echo = FALSE}
### Get trust data on people with mental health conditions ###
female_male_trust <- subset(female_male_data, Section == "Feels the healthcare system can be trusted")

trust <- ggplot(female_male_trust, aes(x = Modules, y = `Total persons`, group = Gender, fill = Gender)) +
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_text(aes(label=`Total persons`), vjust=1.6, color="black",
            position = position_dodge(0.9), size=3.5)+
  theme_minimal() + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 25)) +
  ylab("Total Population")

trust + scale_fill_manual(values=c('#d9848e','#84a8d9'))

```


## Socioeconomic Status

```{r discrimination, fig.cap="Has experienced discrimination in last 12 months", echo = FALSE}
### Get discrimination data ###
person_discrim <- subset(persons_data, Modules == "Has experienced discrimination in last 12 months", select = -c(Section, Modules))

### Transpose dataframe ###
person_discrim <- person_discrim |>
  tibble::rownames_to_column() |>
  pivot_longer(-rowname) |>
  pivot_wider(names_from=rowname, values_from=value) |>
  rename(
    Category = name,
    Population = `1`
  )

### Graph discrimination data ###
person_discrim |>
  ggplot(aes(x = `Category`, y = Population)) +
  geom_bar(stat="identity", position=position_dodge()) + 
  geom_text(aes(label = Population), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5)+
  theme_minimal() + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 15))

```

```{r fig.height = 8, fig.width = 10, fig.cap="Mental Health Conditions vs. Level of highest non-school qualification", echo = FALSE}
### Get bachelor and discrimination data ###
persons_edu_bach <- subset(mutate_all(persons_data, as.character), Modules == "Bachelor Degree or above", select = -c(Section))
### Transpose ###
persons_edu_bach <- persons_edu_bach |>
  tibble::rownames_to_column() |>  
  pivot_longer(-rowname) |>
  pivot_wider(names_from=rowname, values_from=value) |>
  mutate("Qualification" = "Bachelor Degree or above") |>
  slice(-1)

### Get diploma and discrimination data ###
persons_edu_dip <- subset(mutate_all(persons_data, as.character), Modules == "Advanced Diploma/Diploma or below", select = -c(Section))
### Transpose ###
persons_edu_dip <- persons_edu_dip |>
  tibble::rownames_to_column() |>  
  pivot_longer(-rowname) |>
  pivot_wider(names_from=rowname, values_from=value) |>
  mutate("Qualification" = "Advanced Diploma/Diploma or below") |>
  slice(-1)

### Get no non-school and discrimination data ###
persons_edu_none <- subset(mutate_all(persons_data, as.character), Modules == "No non-school qualification", select = -c(Section))
### Transpose ###
persons_edu_none <- persons_edu_none |>
  tibble::rownames_to_column() |>  
  pivot_longer(-rowname) |>
  pivot_wider(names_from=rowname, values_from=value) |>
  mutate("Qualification" = "No non-school qualification") |>
  slice(-1)

### Merge all education and discrimination data and rename columns###
persons_edu <- merge (x = persons_edu_none, 
                      y = merge(x = persons_edu_bach, y = persons_edu_dip, all = TRUE), 
                      all = TRUE
                      ) %>%
  rename (
    Condition = 1,
    Population = 2
  )

persons_edu <- transform(persons_edu, Population = as.numeric(Population))


### Graph discrimination data ###
persons_edu |>
  ggplot(aes(x = Condition, y = Population, group = Qualification, fill = Qualification)) +
  coord_flip() + 
  geom_col(stat="identity", width = 0.8, position=position_dodge(0.8)) + 
  geom_text(aes(label = Population), hjust = 0, vjust = 0.4, color="black",
            position = position_dodge(0.8), size=4)+
  theme_minimal() + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 30)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "bottom") +
  guides(fill=guide_legend(nrow=3, byrow=TRUE)) +
  xlab("Mental Health Condition") +
  labs(fill='Highest Level of Non-School Qualification') 

```

Non-school qualifications are educational attainments other than those of pre-primary, primary or secondary education qualifications.



Talk more about it.

Also bills and their average (Figure \@ref(fig:stressors)). (Notice how you can change the height and width so they don't take the whole page?)

Talk way more about it. 

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @theguardian If we were interested in baseball data then @theguardian could be useful. 

We can use maths by including latex between dollar signs, for instance $\theta$.


## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

## Supplementary Survey



\newpage


# References


